Region-network hierarchical sparsity priors for\\ high-dimensional inference in brain imaging

### 2016-01-27 paper/paper.tex
some math

\paragraph{Problem formulation}

Sparse linear models
encode geometric prior information
topology
local sets of voxels

Group-sparsity is a first step towards the more general idea that
a regularization function can encourage sparse solutions with a particular structure.

it is not realistic to assume that all of the tasks share the same set of rele- vant inputs as in the L1/L2-regularized regression. A subset of highly related outputs may share a common set of relevant inputs, whereas weakly related outputs are less likely to be affected by the same inputs.

structured regularization
We might therefore gain in the quality of the factors
induced by enforcing directly this a priori

groups at multiple granularity

tree-guided group lasso

encourage structured shrinkage effect

l1 = unstructured sparsity-inducing penalty

Our method extends the L1/L2 penalty to the tree-lasso penalty
by letting the hierarchically-defined groups overlap. 
the tree lasso is a special case of overlapping group lasso

for every column u of U, it compute a column v of V solving

we aim at learning a weight vector w ∈ Rp and an intercept b ∈ R
such that the prediction of y can be based on the value of w⊤x + b.

We omit a bias term, since the data were mean-centered
and unit-variance scaled.
The scalar b is not particularly informative

however the vector w corresponds to a volume that
can be represented in brain space as a volume

hierarchial tree = more generally into a directed acyclic graph

more precisely, we denote by X ∈ Rn×p the design matrix
assembled from n fMRI volumes and by y ∈ Rn the corresponding n targets.
In other words, each row of X is a p-dimensional sample,
i.e., an activation map of p voxels related to one stimulus presentation.
for visualization of the predictive pattern of voxels.

Learning the parameters (w, b) remains challenging
since the number of features (104 to 105 voxels) exceeds
by far the number of samples (a few hundreds of volumes).

The scalar b is not particularly informative,
however the vector w corresponds to a volume that
can be represented in brain space as a volume
for visualization of the predictive pattern of voxels.

each row of X is a p-dimensional sample,
i.e., an activation map of p voxels related to one stimulus presentation.

To address this issue, dimensionality reduction attempts to
find a low dimensional subspace that concentrates
as much of the predictive power of
the original set as possible for the problem at hand.
-> we do not want to do preliminary feature selection or
dimensionality reduction
or feature agglomeration because we want to fit one model parameter
to each brain voxel for maximal interpretability
This corresponds to discarding some columns of X.

The essential shortcoming of the Elastic net is that
it does not take into account the spatial structure of the data,
which is crucial in this context

Craddock clusters are often used for feature agglomeration
into parcels
-> exploits only a part of the data

dual-level spatial structure
sparse hierarchical regularization
structured sparsity-inducing regularization
the root of the tree T is the unique cluster that gathers all the voxels,

It is a generalization of the traditional $\ell_1$-norm
$\Omega(\mathbf{w}) = \sum_{j=1}^p|\mathbf{w}_j|$
ignores structure

\cite{jenatton2011multi}

\textit{structured sparsity}

\cite{huang2011learning, morales2010family, jenatton2011structured}

a node $j$ of $\mathcal{T}$,
we denote by $g_j \subseteq \{1,..,q\}$ the set of indices that record
all the descendants of $j$ in $\mathcal{T}$

the family of sparsity-inducing norms has recently been extended
by hierarchical sparsity penalty terms
\cite{zhao2009composite}.

\begin{equation}
  \Omega(\mathbf{w}) = \sum_{g \in G}||\mathbf{w}_g||_2 = \sum_{g \in G}\sqrt{\sum_{j \in g}\mathbf{w}_j^2}
\end{equation}

For example, when G is the set of all singletons, Ω is the
usual l1 norm (assuming that all the weights are equal to 1).

l1/l2 mixed norm is convex

Discarding coefficients belonging to a network group will naturally enforce
discarding the coefficients belonging to each of its descendent region groups.
Conversely,
variable selection of a network group will also enforce
selection of all voxel of its descendent group regions.
Single region groups can however be set to zero (unselected)
or non-zero (selected)
without analogous effect on the parent network group.

At the between-group level,...
$\Omega$ exerts $\ell_1$-like variable selection on
the $(||\mathbf{w}_g||_2)_{g\in G}$ groups,
yielding a maximum of $g \in G$ to be zeroed out
\cite{jenatton2011structured}.
The important consequence is that also all descentes of such a zeroed
group $g \in G$ will be descarded.
Conversely,
if one group $g$ is selected,
then all the ancestral groups will also be selected.
Thus, statistical estimation will be improved by enticing
entire voxel sets to be selected or discarded as predictive, 
although one individual coefficient is computed for each voxel.

-> it is a $(\ell_1, \ell_2)$-mixed norm
-> between-group sparsity effect by l1
-> within-group shrinkage effect by l2

\begin{equation}
  \Omega(\mathbf{w}) = \sum_{g \in G} \eta_g ||\mathbf{w}_g||_2
\end{equation}

$(\eta_g)_{g \in G}$ are positive weights for the groups

fit to the data is measured through
a convex loss function (w, b) 􏰀→ L(y, X, w, b) ∈ R+.

\paragraph{Classification}

logistic loss function

\begin{equation}
  P(y=k|x, W, b) = \frac{exp\{x^Tw^k + b_k\}}{\sum_{m=1}^cexp\{x^Tw^m + b_m\}}
  \argmin \frac{1}{2}||u-v||_2^2 + \lambda\Omega(\mathbf{w})
\end{equation}

$\lambda > 0$

bias is omitted because $\mathbf{X} and \mathbf{y}$
are mean-centered and unit-variance scaled.

In this setting, and given a new fMRI volume x,
we make predictions by choosing the label that maximizes
the class-conditional probabil- ities (3.1), that is, argmaxk∈{1,...,c}Prob(y = k|x; W∗, b∗)

One-versus-rest scheme

\paragraph{Regression}

squared error as loss

\begin{equation}
  \argmin \frac{1}{2}||\mathbf{y - Xw}||_2^2 + \lambda\Omega(\mathbf{w})
\end{equation}

$\lambda > 0$

Prediction for a new fMRI volume x is then simply performed by computing
the dot product x⊤w∗

\paragraph{Numerical optimization}

Difficult because high-dimensional setting

empirical risk minimization was performed by

The intercept $b$ is left unregularized

>>>>

\paragraph{Problem formulation}
We formulate our estimation problem in the framework of regularized risk
estimation applied to linear models: 
We would like to estimate a good predictor of cognitive task
given a brain image. Let the set \(\mathcal X\subset\R^p\) represent brain
images of \(p > 0\) voxels.

Then we would like to minimize the risk \(\mathcal L(\hat y, y)\), where
\(\hat y = X\hat w + \hat b\), while regularizing to incorporate a useful
prior. Taken together, this can be framed as an optimization problem
\[\argmin_{w, b} \mathcal L(Xw + b, y) + \lambda\Omega(w),\]
where \(\lambda > 0\) and \(\Omega\) is the regularizer.

Brain regions are defined as disjoint groups of voxels. Let \(\mathcal G\)
be a partition of \(\{1, \dots, p\}\), i.e.
\[  \bigcup_{i} g_i = \{1, \dots, p\} \textrm{ and } g_i\cap g_j=\emptyset
\quad\forall i\not=j\]

Brain networks consist of regions and are thus super-regions or groups of 
regions. The set of brain networks \(\mathcal{H}\) is also a partition of 
\(\{1, \dots, p\}\) and in addition it is consistent with \(\mathcal G\) in
the sense that
\[\textrm{for all }g\in\mathcal G, h\in\mathcal H,\quad
\textrm{ either } g\subset h\textrm{ or }g\cap h = \emptyset.\]
This allows a clear association of each region \(g\in\mathcal G\) to a 
network \(h\in\mathcal H\) and thus establishes a tree structure (up to 
adding a root node containing all voxels).

For a brain image \(w\in\R^p\) and a group \(g\), the vector 
\(w_g\in\R^{|g|}\) is defined as the restriction of \(w\) to the coordinates
 in \(g\). The structured penalty incorporating network
and region information can then be written as
\[\Omega(w) = \alpha\sum_{h\in\mathcal H}\eta_h\|w_h\|_2 + \beta\sum_{g\in\mathcal G}\eta_g\|w_g\|_2.\]
According to \cite{Zhou_or_who_was_that} we set \(\eta_g = 1/\sqrt{|g|}\) to
account for varying group size. The hierarchy-level-specific factors 
\(\alpha > 0\) and \(\beta > 0\) are used to trade-off region-weighted and
network-weighted models against each other.

The prediction problem at hand is a multiclass classification. We choose to
attack this using one-vs-rest scheme on a binary logistic regression, whose
loss can be written as
\[\sum_{i=1}^n\log(1 + \exp(-y_i\langle x_i, w\rangle)) + \lambda\Omega(w),\]
if \(y\in{-1, 1}\) and with \(x_i\in\R^p\) the training sample brain images.

{\color{red}
MISSING
\begin{itemize}
\item cross-validation scheme
\item analysis of penalty tradeoff (CV scheme for that?)
\item mentioning of control methods in methods. Trace norm, sparse group sparsity. Anything else? ENET?
\end{itemize}
}

%% Sparse linear models
%% encode geometric prior information
%% topology
%% local sets of voxels
%% 
%% Group-sparsity is a first step towards the more general idea that
%% a regularization function can encourage sparse solutions with a particular structure. 
%% 
%% it is not realistic to assume that all of the tasks share the same set of rele- vant inputs as in the L1/L2-regularized regression. A subset of highly related outputs may share a common set of relevant inputs, whereas weakly related outputs are less likely to be affected by the same inputs.
%% 
%% structured regularization
%% We might therefore gain in the quality of the factors
%% induced by enforcing directly this a priori 
%% 
%% groups at multiple granularity
%% 
%% tree-guided group lasso
%% 
%% encourage structured shrinkage effect 
%% 
%% l1 = unstructured sparsity-inducing penalty
%% 
%% Our method extends the L1/L2 penalty to the tree-lasso penalty
%% by letting the hierarchically-defined groups overlap. 
%% the tree lasso is a special case of overlapping group lasso
%% 
%% for every column u of U, it compute a column v of V solving
%% 
%% we aim at learning a weight vector w ∈ Rp and an intercept b ∈ R
%% such that the prediction of y can be based on the value of w⊤x + b.
%% 
%% We omit a bias term, since the data were mean-centered
%% and unit-variance scaled.
%% The scalar b is not particularly informative
%% 
%% however the vector w corresponds to a volume that
%% can be represented in brain space as a volume
%% 
%% hierarchial tree = more generally into a directed acyclic graph
%% 
%% more precisely, we denote by X ∈ Rn×p the design matrix
%% assembled from n fMRI volumes and by y ∈ Rn the corresponding n targets.
%% In other words, each row of X is a p-dimensional sample,
%% i.e., an activation map of p voxels related to one stimulus presentation.
%% for visualization of the predictive pattern of voxels. 
%% 
%% Learning the parameters (w, b) remains challenging
%% since the number of features (104 to 105 voxels) exceeds
%% by far the number of samples (a few hundreds of volumes). 
%% 
%% The scalar b is not particularly informative,
%% however the vector w corresponds to a volume that
%% can be represented in brain space as a volume
%% for visualization of the predictive pattern of voxels.
%% 
%% each row of X is a p-dimensional sample,
%% i.e., an activation map of p voxels related to one stimulus presentation.
%% 
%% To address this issue, dimensionality reduction attempts to
%% find a low dimensional subspace that concentrates
%% as much of the predictive power of
%% the original set as possible for the problem at hand.
%% -> we do not want to do preliminary feature selection or
%% dimensionality reduction
%% or feature agglomeration because we want to fit one model parameter
%% to each brain voxel for maximal interpretability
%% This corresponds to discarding some columns of X.
%% 
%% The essential shortcoming of the Elastic net is that
%% it does not take into account the spatial structure of the data,
%% which is crucial in this context
%% 
%% Craddock clusters are often used for feature agglomeration
%% into parcels
%% -> exploits only a part of the data
%% 
%% dual-level spatial structure
%% sparse hierarchical regularization
%% structured sparsity-inducing regularization
%% the root of the tree T is the unique cluster that gathers all the voxels,
%% 
%% It is a generalization of the traditional $\ell_1$-norm
%% $\Omega(\mathbf{w}) = \sum_{j=1}^p|\mathbf{w}_j|$
%% ignores structure
%% 
%% 
%% \cite{jenatton2011multi}
%% 
%% \textit{structured sparsity}
%% 
%% \cite{huang2011learning, morales2010family, jenatton2011structured}
%% 
%% 
%% a node $j$ of $\mathcal{T}$,
%% we denote by $g_j \subseteq \{1,..,q\}$ the set of indices that record
%% all the descendants of $j$ in $\mathcal{T}$
%% 
%% the family of sparsity-inducing norms has recently been extended
%% by hierarchical sparsity penalty terms
%% \cite{zhao2009composite}.
%% 
%% 
%% \begin{equation}
%%   \Omega(\mathbf{w}) = \sum_{g \in G}||\mathbf{w}_g||_2 = \sum_{g \in G}\sqrt{\sum_{j \in g}\mathbf{w}_j^2}
%% \end{equation}
%% 
%% For example, when G is the set of all singletons, Ω is the
%% usual l1 norm (assuming that all the weights are equal to 1).
%% 
%% l1/l2 mixed norm is convex
%% 
%% 
%% Discarding coefficients belonging to a network group will naturally enforce
%% discarding the coefficients belonging to each of its descendent region groups.
%% Conversely,
%% variable selection of a network group will also enforce
%% selection of all voxel of its descendent group regions.
%% Single region groups can however be set to zero (unselected)
%% or non-zero (selected)
%% without analogous effect on the parent network group.
%% 
%% 
%% 
%% 
%% At the between-group level,...
%% $\Omega$ exerts $\ell_1$-like variable selection on
%% the $(||\mathbf{w}_g||_2)_{g\in G}$ groups,
%% yielding a maximum of $g \in G$ to be zeroed out
%% \cite{jenatton2011structured}.
%% The important consequence is that also all descentes of such a zeroed
%% group $g \in G$ will be descarded.
%% Conversely,
%% if one group $g$ is selected,
%% then all the ancestral groups will also be selected.
%% Thus, statistical estimation will be improved by enticing
%% entire voxel sets to be selected or discarded as predictive, 
%% although one individual coefficient is computed for each voxel.
%% 
%% -> it is a $(\ell_1, \ell_2)$-mixed norm
%% -> between-group sparsity effect by l1
%% -> within-group shrinkage effect by l2
%% 
%% 
%% \begin{equation}
%%   \Omega(\mathbf{w}) = \sum_{g \in G} \eta_g ||\mathbf{w}_g||_2
%% \end{equation}
%% 
%% $(\eta_g)_{g \in G}$ are positive weights for the groups
%% 
%% 
%% fit to the data is measured through
%% a convex loss function (w, b) 􏰀→ L(y, X, w, b) ∈ R+. 
%% 
%% 
%% \paragraph{Classification}
%% 
%% logistic loss function
%% 
%% \begin{equation}
%%   P(y=k|x, W, b) = \frac{exp\{x^Tw^k + b_k\}}{\sum_{m=1}^cexp\{x^Tw^m + b_m\}}
%%   \argmin \frac{1}{2}||u-v||_2^2 + \lambda\Omega(\mathbf{w})
%% \end{equation}
%% 
%% $\lambda > 0$
%% 
%% bias is omitted because $\mathbf{X} and \mathbf{y}$
%% are mean-centered and unit-variance scaled.
%% 
%% In this setting, and given a new fMRI volume x,
%% we make predictions by choosing the label that maximizes
%% the class-conditional probabil- ities (3.1), that is, argmaxk∈{1,...,c}Prob(y = k|x; W∗, b∗)
%% 
%% One-versus-rest scheme
%% 
%% 
%% \paragraph{Regression}
%% 
%% squared error as loss
%% 
%% \begin{equation}
%%   \argmin \frac{1}{2}||\mathbf{y - Xw}||_2^2 + \lambda\Omega(\mathbf{w})
%% \end{equation}
%% 
%% $\lambda > 0$
%% 
%% Prediction for a new fMRI volume x is then simply performed by computing
%% the dot product x⊤w∗
%% 
%% \paragraph{Numerical optimization}
%% 
%% Difficult because high-dimensional setting
%% 
%% empirical risk minimization was performed by
%% 
%% 
%% The intercept $b$ is left unregularized
